---
title: "Thinning strategies in *E. pellita*"
author: "Araujo MJ"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: '2'
    df_print: paged
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    number_sections: yes
    fig_caption: yes
    toc-location: left
  html_notebook:
    code_folding: hide
    toc: yes
    toc_depth: '2'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    number_sections: yes
    fig_caption: yes
    toc-location: left
---
```{r, include=FALSE}
knitr::opts_chunk$set(
  echo = T,
  message = F,
  warning = F
)
```

# Wrangling dataset

```{r inst.pct}
# Install packages
## devtools::install_github('famuvie/breedR') # --> Fail to install
# The codes bellow succeed to install
## source("http://famuvie.github.io/breedR/src/setup_repo.R")
## install.packages('breedR')
```


```{r pct, message=F, warning=F}
library(tidyverse)
library(openxlsx)
library(breedR)
```

```{r load dt, message=F, warning=F}
Data <- read.xlsx("Data/Data.xlsx", sheet = "Progenie") |>
  # Excluding Treatment = 1
  filter(is.na(Tratamento))
knitr::kable(head(Data, 10))
```

```{r load func}
# Load `DiagFunc` and 'others' functions
source(
  "https://raw.githubusercontent.com/AraujoMJ/AraujoMJ_Functions/main/Functions_mja_2022_v1.R"
)
# Load `Thinning_BreedR` function
source(
  "https://raw.githubusercontent.com/AraujoMJ/AraujoMJ_Functions/main/Thinning_BreedR.R"
)
```

# Diagnostic analysis {.tabset}

## ***DAP:***

```{r DiagFuncDAP, message=F, warning=F, fig.height=6, fig.width=8}
# names(Data)
par(mfrow = c(2, 2))
DiagDAP <-
  DiagFunc(
    Rep = "Bloco",
    Trait = "DAP.3",
    Trat = "Genotipo",
    data1 = Data,
    Exp_Name = " ",
    Exp = NULL,
    plot_diag1 = T,
    plot_diag2 = T,
    nDiag = 1,
    verbose = F,
    ColumnNames_To_Return = c(
      "Bloco",
      "Genotipo",
      "Procedencia",
      "especie_id",
      "Arv",
      "CONCAT.",
      "DAP.1",
      "Alt.1",
      "Vol.1"
    )
  )
# ~ 
```

## ***ALT:***

```{r DiagFuncALT, message=F, warning=F, fig.height=6, fig.width=8}
# names(Data)
par(mfrow = c(2, 2))
DiagALT <-
  DiagFunc(
    Rep = "Bloco",
    Trait = "Alt.3",
    Trat = "Genotipo",
    data1 = Data,
    Exp_Name = " ",
    Exp = NULL,
    plot_diag1 = T,
    plot_diag2 = T,
    nDiag = 1,
    verbose = F,
    ColumnNames_To_Return = c(
      "Bloco",
      "Genotipo",
      "Procedencia",
      "especie_id",
      "Arv",
      "CONCAT.",
      "DAP.1",
      "Alt.1",
      "Vol.1"
    )
  )
#DiagALT
```

## ***Vol:***

```{r DiagFunc Vol, message=F, warning=F, fig.height=6, fig.width=8}
# names(Data)
par(mfrow = c(2, 2))
DiagVOL <-
  DiagFunc(
    Rep = "Bloco",
    Trait = "Vol.3",
    Trat = "Genotipo",
    data1 = Data,
    Exp_Name = " ",
    Exp = NULL,
    plot_diag1 = T,
    plot_diag2 = T,
    nDiag = 1,
    verbose = F,
    ColumnNames_To_Return = c(
      "Bloco",
      "Genotipo",
      "Procedencia",
      "especie_id",
      "Arv",
      "CONCAT.",
      "DAP.1",
      "Alt.1",
      "Vol.1"
    )
  )
#
```

```{r mergeData}
# Merge data
DataALL <- left_join(
  left_join(
    # DAP and ALT
    DiagDAP$DataAfterDiag,
    DiagALT$DataAfterDiag,
    by = c("Bloco",
           "Genotipo",
           "Procedencia",
           "especie_id",
           "Arv",
           "CONCAT.",
           "DAP.1",
           "Alt.1",
           "Vol.1")
  ),
  DiagVOL$DataAfterDiag,
  by = c("Bloco",
         "Genotipo",
         "Procedencia",
         "especie_id",
         "Arv",
         "CONCAT.",
         "DAP.1",
         "Alt.1",
         "Vol.1")
)

knitr::kable(head(DataALL, 10))
```

# Genetic analysis

## Build Pedigree

### Recode Genotipo

```{r prep ped}
# Map genotype
RecodeGen <- DataALL |>
  filter(Procedencia != "Clone") |>
  distinct(Genotipo) %>%
  mutate(recoded_genotype = seq_len(nrow(.)))

# Insert column with recode Genotipo: mother
DataALL$mother <-
  RecodeGen[match(DataALL$Genotipo, RecodeGen$Genotipo), "recoded_genotype"] |>
  as.numeric()

# Creating dataset without Clone
Progenies <- DataALL |>
  filter(Procedencia != "Clone") |>
  mutate(ID = as.numeric(seq_along(Genotipo) + 1000),
         father = as.numeric(NA),
         Plot = interaction(Bloco, Genotipo, sep = ":"))

```

### Prepare pedigree: `build_pedigree` function from breedR package

```{r pedBreedR}
Pedigree <- build_pedigree(
  self = "ID",
  sire = "father",
  dam = "mother",
  data = Progenies
)
knitr::kable(tail(as.data.frame(Pedigree), 10))
```

## Adjust model: `remlf90` function from breedR package

```{r mod_1, message=F, warning=F}
system.time(modDAP <- remlf90(
  DAP.3 ~ 1 + Bloco,
  random = ~ Plot + Procedencia,
  genetic = list(
    model = "add_animal",
    pedigree = Pedigree,
    id = "ID"
  ),
  data = Progenies |>
    mutate(Bloco = as.factor(Bloco),
           Procedencia = as.factor(Procedencia),
           Plot = as.factor(Plot)),
  method = "em",
  progsf90.options = 'EM-REML 10'
  )
)
summary(modDAP)
```

## Deviance Analysis

***By using*** **`Deviance_BreedR` function**

```{r applyDeviance, message=F, warning=F}
DevianceAnalyses <- Deviance_BreedR(
  Trait = "DAP.3",
  Model = modDAP,
  Data = Progenies,
  Pedigree = Pedigree,
  Method = "EM",
  model_type = "std_animal"
)
DevianceAnalyses
```

## Extract estimates

```{r Varcomp}
Extract_h2a_breedR(Model = modDAP, 
            random_effect = c("Plot", "Procedencia"),
            model_type = "std")
```

# Extract Blups

***By using*** `BV_BreedR` ***function***

-   **BV** from provenances, families and trees:

```{r BVfam}
# names(Progenies)
NamesReturn <-
  c(
    "ID",
    "Bloco",
    "Genotipo",
    "Procedencia",
    "especie_id",
    "Arv",
    "CONCAT.",
    "DAP.1",
    "Alt.1",
    "Vol.1",
    "DAP.3",
    "Alt.3",
    "Vol.3",
    "mother"
  )
BV <- BV_BreedR(
  data = Progenies,
  Model = modDAP,
  NamesID = "ID",
  Pedigree = Pedigree,
  ColumnFamilies = "mother",
  Rank = T,
  NamesReturnDataset = NamesReturn,
  model_type = "std",
  random_effect = "Procedencia",
  fixed_effect_av = "Bloco"
)

```

## Organizing breeding values on datasets

```{r BV_org}
# BV from trees
Data_BV <- BV$Data_Total
  # Input provenance effects
Data_BV$BV_Procedencia <-
  BV$BV_random_effect[match(Data_BV$Procedencia, rownames(BV$BV_random_effect)), "value"]
  # Input families effects
Data_BV$BV_family <-
  BV$BV_Parents[match(Data_BV$mother, BV$BV_Parents$Family), "a"]

# BV_from_families
BV_Family <- BV$BV_Parents
  # Input family information: 'Genotipo' column on original data
BV_Family$Genotipo <-
  Data_BV[match(BV$BV_Parents$Family, Data_BV$mother), "Genotipo"]

  # Input provenance information
BV_Family$Procedencia <-
  Data_BV[match(BV$BV_Parents$Family, Data_BV$mother), "Procedencia"]
  # Input provenance effects
BV_Family$a_prov <-
  BV$BV_random_effect[match(BV_Family$Procedencia, rownames(BV$BV_random_effect)), "value"]

  # Input BV_Total: BV from families + BV from provenances
BV_Family <- BV_Family |>
  rename(a_fam = a,
         s.e_a_fam = s.e,
         `u+a_fam` = `u+a`) |>
  mutate(a_total = a_fam + a_prov,
         BV_Total = `u+a_fam` + a_prov) |>
  dplyr::select(Family, Genotipo, Procedencia, s.e_a_fam,`u+a_fam`, a_fam, a_prov, a_total, BV_Total)

knitr::kable(BV_Family)
```

```{r exportProced_excel, eval=FALSE}
write.xlsx(BV$BV_random_effect |>
             rownames_to_column(var = "Provenance"), 
           file = paste0("Output/Blups_proced_prog_test_pellita", Sys.Date(), ".xlsx"))
```

*Breeding values from families (`a_fam`) vs Breeding values from families + provenances (`a_fam` + `a_prov`)*

```{r BV_id_Total}

plot(BV_Family$a_fam ~ BV_Family$a_total)


```

# Running model without provenance effect in the model

## Adjust model

```{r ModelNoProv, message=F, warning=F}
system.time(modDAP2 <- remlf90(
  DAP.3 ~ 1 + Bloco,
  random = ~ Plot,
  genetic = list(
    model = "add_animal",
    pedigree = Pedigree,
    id = "ID"
  ),
  data = Progenies |>
    mutate(Bloco = as.factor(Bloco),
           Plot = as.factor(Plot)),
  method = "em",
  progsf90.options = 'EM-REML 10'
  )
)
summary(modDAP2)

```

## Extract BLUP

```{r BLUP2}

BV2 <- BV_BreedR(
  data = Progenies,
  Model = modDAP2,
  NamesID = "ID",
  Pedigree = Pedigree,
  ColumnFamilies = "mother",
  Rank = T,
  NamesReturnDataset = NamesReturn,
  model_type = "std",
  random_effect = NULL,
  fixed_effect_av = "Bloco"
)
# BLUP from families
BV_Family2 <- BV2$BV_Parents
knitr::kable(BV_Family2)
```

1.  *Breeding values from families on model with provenances effects (`a_fam`) vs Breeding values from families on model without provenances effects (`a`)*

2.  *Breeding values from families + provenance on model with provenances effects (`a_total`) vs Breeding values from families on model without provenance effects (`a`)*

```{r comparisomModels}
# Case 1
plot(BV_Family$a_fam, BV_Family2$a)
# Case 2
plot(BV_Family$a_total, BV_Family2$a)
```

# Thinning Strategies

```{r thinningPellita}
system.time(
  Thinning_pellita <- Thinning_BreedR(
    BV_Column = "a_total",
    Trait = "DAP.3",
    BV_fam = BV_Family,
    Data_Total = BV$Data_Total,
    Family_Data_Total = "mother",
    Bloc_Column = "Bloco",
    nGroups = 4,
    label.group.y = c(1, 1.5, 1, 1),
    Plot.Rank = TRUE,
    save_plot_rank = TRUE,
    IS = NULL,
    id = "ID",
    length_seq_combinations = 1
  )
)
# ~
```

